---
title: "Creating_New_Templates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating_New_Template}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(dsreportr)
```

There are a few tools to make creating derivative packages that incorporate
new templates. As an example, let's consider creating a new "bbireportr" 
package. 

## Usethis inspired approach
The easiest way to create a new template is not to, but rather just modify the YAML header from the skeleton markdown in a new file. That's great for a one-off. If that doesn't work, create a new package (based on devtools, at least in rstudio). Then, from the project directory:
```
dsreportr::create_template_in_package("bbireportr::bbi_pdf")
```

This will create the necessary files, etc to make a new template. Note that the function provides copious output during the creation.

What if you want a banner image with the template? No problem, you can run:
```
dsreportr::create_template_in_package("bbireportr::bbi_pdf", "image.jpg")
```

If you think about adding a banner after the fact (after creating the template), you 
can just use:

```
dsreportr::use_banner("bbireportr::bbi_pdf", "image.jpg")
```
 

## Manual Steps

You can also manually create a template for use. The basic components of the package that are required include the following.

### bbi_pdf.R

`R/bbi_pdf.R` is a wrapper around `bookdown::pdf_document2()` that provides
a way to insert the latex template. If this isn't clear, don't worry since there
isn't much that is needed. Since the template comes from this package (dsreportr), there is no need to do anything except name the function. Here is an example of what it could be:

```
bbi_pdf <- dsreportr::ds_pdf
```

However, if you want to use a roxygen2 template and use rstudio `Code->Insert Roxygen Skeleton`, then you would create a bit more:
```
bbi_pdf <- function(...) {dsreportr::ds_pdf(...)}
```

That's it. We can now specify in a YAML markdown report header:

```
output:
  bbireportr::bbi_pdf()
```

### Template
Next, we create the following directory: `inst/rmarkdown/templates/bbi_pdf`. 
This directory will hold the banner image (if you want one), the skeleton
markdown and the description of the template.

- The banner image is easy - put a jpg/png file in the `inst/rmarkdown/templates/bbi_pdf/resources` directory. You want to make sure it is appropriately sized (full width looks nice, but other options are reasonable). 

- A description of the template should be in `inst/rmarkdown/templates/bbi_pdf/template.yaml`. This is a simple file with only a couple of lines:
```
name: Biostatistics, Bioinformatics and Informatics PDF Report Output
description: A Biostatistics, Bioinformatics and Informatics PDF report.
create_dir: false
```
You may want more inspired text, but that is sufficient. Note that the name will show up in the `File->New File->R Markdown->Templates` selection in Rstudio.

- A skeleton markdown file, providing the necessary components to run the template. Note that this package provides a complete `skeleton.Rmd` file which can be copied to: `inst/rmarkdown/templates/bbi_pdf/skeleton`. You don't have 
to find the skeleton file, though. You could execute the following:
```
dsreportr::new_markdown("inst/rmarkdown/templates/bbi_pdf/skeleton/skeleton.Rmd")
```

From this file, you'll want to change two fields:

```
banner: "`` `r
dsreportr::banner("bbireportr::bbi_pdf")` ``"
output:
  bbireportr::bbi_pdf:
    latex_engine: pdflatex
    keep_tex: false
```

That's it. At this point, you should have a workable package that implements a new R markdown template.
